One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction

### 2017-03-20 Zhiyu.tex
Jan26 PRE format, tex

In mean-field perspective, the physics of atom gas in Bose-Einstein Condensation can be described by the Gross-Pitaevskii equation\cite{Dalfovo1999}\cite{Stringari1996}, where the breathing frequency for 1D Bose gas is calculated to be $\Omega_{GP}=\sqrt{3}\omega_0$. Meanwhile, the experiment of Ref.\cite{Haller2009} has found a regime of interactions where the breathing-mode frequency approaches the Gross-Pitaevskii prediction. However, since mean-field approach which generally omit quantum fluctuation is successful in this problem, one may wonder how quantum effect plays the role in the breathing mode. Thus, a natural question would be asked is how breathing frequency is dependent on interaction for classical gas. Though generated in BEC, the breathing mode phenomenon is not restricted to BEC. For a classical gas, the breathing mode still can be excited with a quench. A good knowledge of classical version may help us understand the quantum question better. Surprisingly, to the best of our knowledge, all the literature focus on the breathing mode in BEC, without a careful study of their classical counterpart, which actually seems easier. Since this classical version question has not been answered yet, more study is needed.

One the other hand, in a nonquilibrium many-body system, a more intriguing question is about its thermalization or relaxation property. Thermalization is among the most fundamental process for a many-body interacting system. All systems do not thermalize. An example is the Fermi-Pasta-Ulam paradox [ ] which shows confinement in phase-space so that the system stay non-thermalized for a long time. A lot of research on thermalization property has been done in other classical system. Ref.\cite{Tsuchiya2000}\cite{Yawn1997} studied the one dimensional gravitational system with Lyapunov exponents. Ref.\cite{Jin2013} studied the ring of harmonic oscillators and magnetic moment by examine the canonical distribution of a subsystem . In our paper, we will also probe the system's thermalization time scales from the perspective of canonical distribution as well as Lyapunov exponent. With these approaches, we will show that the thermalization time scale manifest a competition between total energy and interaction energy, so there is not any singular behavior in the thermalization time scales except the trivial divergence at zero interaction.

This paper is organized as follows: In Section \ref{section:preparation}, we perform a scaling to reduce the parameter of the system. In addition, we briefly observe the ground state and reveal two dynamically distinct regime of the system, i.e. strong interaction regime and weak interaction regime. In Section \ref{section:breathing frequency}, we evaluate the breathing frequency according to its mechanism, and compare that with our numerical simulation result. Besides, we also introduce a rotating frame method of treating this problem, which is a convenient tool to analyze simple harmonic systems. In section \ref{section:Thermalization}, we focus on the thermalization property. We first examine the canonical distribution of energy to find out the threshold of thermalization, i.e. under what parameter the thermalization time scale is very large. Then we point out that there are two kind of thermalization (or relaxation) behavior that could be measured in this system. In the end of this section, we measure the Lyapunov Exponent to quantify the time scales and confirm the former discussion.

\section{Preparation}\label{section:preparation}
\subsection{Scaling behaviour}
We investigated the classical gas in harmonic trap with a repulsive interaction. We chose one of the simplest types of interaction:
\begin{equation}
F=\left\lbrace
\begin{aligned}
&F=F_0 	&,x\textless\sigma\\
&F=0 	&,x\textgreater\sigma
\end{aligned}
\right.
\label{eq:preparation1}
\end{equation}
so the Hamiltonian is
\begin{equation}
H=\frac{1}{2}m\omega_0^2\sum_{i}x_i^2+\frac{1}{2}m\sum_i v_i^2+\sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}
It can be shown that Eq.\ref{eq:preparation2} can be rewrite into the following form:
\begin{equation}
\tilde{H}=\frac{1}{2}\sum_{i}\tilde{x_i}^2+\frac{1}{2}\sum_i\left(\frac{d\tilde{x}}{d\tilde{t}}\right)^2+\sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless 1}F_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)
\end{equation}
with the transformation:
\begin{equation}
\left\lbrace
\begin{aligned}
&\tilde{H}&=&\frac{H}{m\omega_0^2\sigma^2}\\
&\tilde{x_i}&=&\frac{x_i}{\sigma}\\
&\tilde{t}&=&\omega_0t\\
&\tilde{F_0}&=&\frac{F_0}{m\omega_0^2\sigma}\\
\end{aligned}
\right.
\end{equation}

In this manner, we reduced number of the parameters in our model into three: $\tilde{H}$, $\tilde{F_0}$, $N$. Equivalently, we set $\sigma$, $\omega_0$ and $m$ to 1 in our numerical simulation. Throughout, we will use ``E" and ``$F_0$" to denote their reduced version  $\tilde{H}$, $\tilde{F_0}$.

@@

\subsubsection{Ground states}
For the ground state of this system, one could predict that \\(1) When the $\sigma$ is larger than the size of the cloud (in view of the scaling behavior, large $\sigma$ is equivalent to small $F_0$ limit when we fix $\sigma=1$), it can be easily shown that the ground state is crystal-like, since all every particle keeps an equal distance with its nearest neighbors, which is $\frac{2F_0}{\omega_0^2}$. The low-lying excitation is particles' independent harmonic oscillation around their equilibrium position.\\(2) When $F_0$ is very large the molecules act like a hardcore gas, while the size of particle is just $\sigma$.

In the fig.\ref{fig:GS1}, we can see this crossover from the ``solid-like" limit (left) to the ``hardcore gas" limit (right).

@@

\section{Breathing Frequency}\label{section:breathing frequency}
Usually, to see the response of quantum gas to a quench, one begin simulation with the ground state and then suddenly vary some parameter by a little, for instance, $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$. In that case, one will observe the radius R oscillating at a certain frequency, which is usually called ``breathing mode".
However, it is quite different in the classical version. Firstly, the ground state may be meaningless. In quantum case, the reason why we care about ground state and low lying excitations is that BEC occur at low temperature. But in classical case, there is no BEC. So it is not necessary to focus on low lying excitations in classical case. Besides, since classical limit is $\hbar\rightarrow 0$, no matter how low the kT is, $\hbar/kT$ is always zero so that classical behavious of low lying excitations is quite different from the quantum one. Secondly, it is no longer necessary to vary parameters ``by a little" in the quench. In quantum case, the small quench $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$ means the spectrum is shifted slightly, thus one may expect to see some beat phenomenon({\color{red}{?}}). But in classical case, since $\hbar$ vanishes, no matter how small the quench is, one can never recover or proximate the quantum case. Thus it is not necessary to limit ourselves to small quench.

Furthermore, notice that X and P (normalized) are symmetric for simple harmonic oscillator. In weak interaction limit, the distribution cloud should be circular when the gas is completely thermalized. The difference between thermalized cloud before and after quench is that they are circular under different normalization of phase-space. If we look at both of them in the phase-space normalized according to the final $\omega_0$, we will find shape of the cloud before the quench is an ellipse while the final one is a circle.

Since the number of particles we use in the Molecular Dynamics simulation is limited (usually 5 to 20), the random noise is significant. If we start with a state whose configuration in phase-space is an ellipse slightly deviated from circle, the oscillating amplitude will be too small to be distinguished from noise. Therefore, we start from an extreme case -- line distribution (random x, zero p) and observe the oscillation of R(t). 
\begin{figure}[hbtp]
\center
\includegraphics[scale=0.32]{ZhiyuPictures/freq_scanF_differentE_log_2_with_oscillation_demo.png}
\caption{breathing mode frequency measured at different E and $F_0$. Insert: a demonstration of oscillation, $E_{trap}$ is the total potential energy of particles in the trap, which is proportional to $\sum_{is}{x_i^2}$, thus manifest the oscillation behavior of radius R(t) defined in Eq.\ref{eq:def_of_R}. The breathing frequency is measured by taking Fourier transformation of R(t).}
\label{fig:Breathingfrequency1}
\end{figure}

@@

\subsection{Rotating frame in Phase-space}
Traditionally, we think of there is a distribution cloud in phase space, and due to harmonic trap, the cloud will rotate at frequency $\omega_0$. When there is interaction, we may think there is a small modification of the frequency $\omega=\omega_0+\delta$. Now we want to understand how $\delta$ is dependent on parameters, so it is better to stand in the rotating frame in the phase space. In this frame, the picture we will see is as follow:\\
\begin{enumerate}[\textbf{*}]
\item All particles are stationary when there is no interaction.
\item The real X and P axis are rotating counter-clockwise, using real X axis to measure distance between particles to determine the interaction at each moment.
\item When there is interaction, particles will gain a ``velocity" in phase space: $\dot{P}=F/m$ . Other effects are all cancelled by frame rotation.
\end{enumerate}

This picture is actually a classical version of the interaction picture. With this picture, we will soon find the dynamic analysis greatly simplified.

In our system, we choose initial velocity to be zero, which means, the particles are aligned on $X|_{t=0}$ axis in phase-space at the beginning. If the breathing frequency of the system is $\omega=\omega_0+\delta$, we will expect to see a line (which may gradually deform into an oval cloud or even an isotropic cloud) rotating at $\delta$ in rotating frame. If we plot x-p with t, we expect to see a spiral motion. Since the motion we see is an additional rotation in rotating frame, we will call it precession. This precession directly leads to the $\delta$.

Fig.\ref{fig:Breathingfrequency2_1} show this precession motion in the rotating frame in phase-space.

\begin{figure*}
\subfigure[$t=0$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=1-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=1-5T_rev.eps}
\end{minipage}
}
\subfigure[$t=\frac{1}{5}T$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=2-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=2-5T_rev.eps}
\end{minipage}
}\subfigure[$t=\frac{2}{5}T$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=3-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=3-5T_rev.eps}
\end{minipage}
}\subfigure[$t=\frac{3}{5}T$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=4-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=4-5T_rev.eps}
\end{minipage}
}\subfigure[$t=\frac{4}{5}T$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=5-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=5-5T_rev.eps}
\end{minipage}
}
\caption{N=50,F0=100,E=50000. Upper(lower) five pictures are snapshot of the configuration of the cloud in the stationary(rotating) frame at t=$\frac{1}{5}T$, $\frac{2}{5}T$, $\frac{3}{5}T$, $\frac{4}{5}T$, $T$ respectively.}
\end{figure*}

@@

(See fig.\ref{fig:Breathingfrequency3}(a))Two particles flip to another line(thick gray line) which deviate a small angle $\alpha$ away from original configuration. Similar process happen when there are more particles. In every period of harmonic oscillation, each particle meet N particles and collide 2N times, while half of the collisions (N times) are between particles with huge difference in momentum, which is just the process shown in fig.\ref{fig:Breathingfrequency3}. (As for the remaining half of collisions, colliding particles have small difference in their momentum. Let us say, both particles have positive momentum. In fig.\ref{fig:Breathingfrequency3}, it can be shown that two particles on the same half of the P-axis cannot have strong effect on the rotation of distribution.)

Let us estimate the precession angle $\alpha$ in one collision by $\frac{\sigma}{R}$, which follows from the two-particle case study. Radius of the cloud R can be estimated according to $E=Nm\omega_0^2R^2$
We will get the precession angular velocity 
\begin{equation}
\delta=2\omega_0\frac{N^\frac{3}{2}m^{\frac{1}{2}}E^{-\frac{1}{2}}\sigma\omega_0}{2\pi}
\label{eq:breathingfrequency1}
\end{equation}

When $F_0$ is not that large, the frequency behavior seems complicated (see fig.\ref{fig:Breathingfrequency1}). Usually, when $F_0$ increase, frequency goes down below 2 first and then rise up and converge to some value higher than 2. The reason could also be well understood with the mechanism described before. In the former discussion, we goes to high $F_0$ limit, which means particles exchange momentum in infinitesimal time. For $F_0$ is not very big, the finite interaction time should be taken into consideration. During this time, the behavior of each particle in phase space is exactly moving along the direction of the real P axis at 'velocity' $F0/m$. Since the real P axis is rotating counter-clockwise, the particle will follow the circular trajectory(orange and green dashed line in fig.\ref{fig:Breathingfrequency3}) the $\alpha$ that we once use to estimate the precession motion should decrease with the increase of interaction time. There should be some ``critical point" where $\alpha$ goes from plus to minus (dashed green line). At this point, we will see the precession stop so that the frequency is 2.

@@

The dependence of frequency on energy and $F_0$ is shown in fig.\ref{fig:Breathingfrequency4}. For the first thing, as expected when $F_0$ is high the frequency behavior is well-described by linear prediction Eq.\ref{eq:breathingfrequency1}. Secondly, the figure shows the frequency's dependence on $F_0$. As we explained, when $F_0$ is not so large, the precession angle decrease. In language of real space, the finite $F_0$ create a finite approaching time, thus consumes the time saved by finite-range collision. As a result, the frequency should be lower when $F_0$ decrease, and reach 2 earlier.

the critical point that frequency falls to 2 can be estimate by:
\begin{equation}
\omega_0 \tau=\alpha
\end{equation}
which states that the approaching time $\tau$ exactly cancels the precession angle derived before
$\tau$ can be estimated by $R/F_0$. With some math, one can show that this condition is equivalent to

\begin{equation}
\frac{E}{N}=F_0\sigma
\end{equation}
which is the same as the criterion of whether two particle will bounce or pass by if one estimate the average internal energy of a pair with $\frac{E}{N}$. This argument indicate the reason why we see frequency stop falling down below 2 as one should expect with previous discussion:\
All our previous analysis based on fig.\ref{fig:Breathingfrequency3} only studies the case of bouncing. If the energy of particles are so large that they often overcome the interaction potential barrier and pass each other, the situation would be completely different. In that case, in fig.\ref{fig:Breathingfrequency3}, particles will move a little bit out of their initial position, and soon move back, almost return to their initial place. In this manner, the interaction do not affect the distribution of the cloud in phase space. Thus, the frequency will stay near 2 when E is large.

Before completing the discussion, it is necessary to point out that none of the argument here require ``energy thermalization". In addition, if the system is completely thermalized in terms of their phase of orbit in phase space, the argument will no longer work -- there is a stable isotropic distribution in phase space, no matter how the cloud rotate, no oscillation can be observed.

@@

We can always think of them in the reference frame of their center of mass ``C". In frame C, one can easily show that when there is no interaction what we will see is that two particles are bound together by a harmonic trap center at C (black line in fig.\ref{fig:thermalization2}). So their relative motion is also a harmonic oscillation.

In this case, the system has two frequency components. The first one is the frequency of C, which is just $\omega_0$ . The other one is the frequency of their relative motion $\omega_r$, which is slightly deviated from $\omega_0$. The deviation grows with the increase of $F_0$ and the decrease of internal energy of the pair. As a result, when the $F_0$ is small and when the E is large, a beat with low frequency $\omega_r-\omega_0=\delta$ will exist. This phenomenon means energy transfer between particles are inefficient, thus it will slow down the thermalization process.

@@

\subsection{Thermalization condition}
The main obstruct to get an thermalized state (which can be examined by observing distribution) is the low frequency oscillation mentioned before. Because once these modes are excited, the relaxing time could be very long($\textgreater 10^4$). As a result, to achieve ergodic state, we have to avoid such low frequency oscillation. According to our discussion before, the solution is to make internal energy of each two-particle pair not ``too large". Though it is impossible to express the internal energy of every pair in terms of the total energy $E$, we can estimate it by the average energy, which is $\frac{E}{N}$. At least they are of the same order. The thermalization condition could be given by:
\begin{equation}
F_0\sigma\sim\frac{E}{N}
\label{eq:thermalizatiton condition}
\end{equation}

@@

As is shown in fig.\ref{fig:thermalization5} above, the critical point for reaching ``Boltzmann-like" distribution (later we will verify that it is Boltzmann distribution) is $F_0=500$-$F_0=1000$ (we always set $\sigma=1$), while the average energy $\sim 1000$, as we expected, they are of the same order. As a supplementary proof of our argument, fig.\ref{fig:thermalization6} and fig.\ref{fig:thermalization7} show the great difference of single-particle energy between two examples of below and above the thermalization threshold ($F_0=200$ $\&$ $F_0=1000$ for $N=10, E=10000$): in the former one there is always some particle maintained at high excitation, while in the latter the system probably goes to ergodic. To date, we have verified that the threshold of thermalization, which is $\frac{NF_0\sigma}{E}\sim1 $ , and corresponding feature of their energy distribution in two side of the threshold.

>>>>

In mean-field perspective, the physics of atom gas in Bose-Einstein Condensation can be described by the Gross-Pitaevskii equation\cite{Dalfovo1999}\cite{Stringari1996}, where the breathing frequency for 1D Bose gas is calculated to be $\Omega_{GP}=\sqrt{3}\omega_0$. Meanwhile, the experiment of Ref.\cite{Haller2009} has found a regime of interactions where the breathing-mode frequency approaches the Gross-Pitaevskii prediction. However, since mean-field approach which generally omit quantum fluctuation is successful in this problem, one may wonder how quantum effects play the role in the breathing mode. Thus, a natural question would be asked is that how breathing frequency is dependent on interaction for classical gas. Though generated in BEC, the breathing mode phenomenon is not restricted to BEC. For a classical gas, the breathing mode still can be excited with a quench. Usually, a good knowledge of the classical version may help us understand the quantum one better. David Guéry-Odelin, Francesca Zambelli, Jean Dalibard, and Sandro Stringari answered this question in ref.\cite{Guery-Odelin1999}. They used the Boltzmann equation as well as molecular dynamics simulation to study the monople mode and quarulple mode of the 2-D classical gas in a harmonic trap. But in that work, the interaction is elastic collision, i.e. zero-range-infinite-strength interaction. We think introducing a finite range for the repulsive interaction will definitely enrich the physics in this system which looks simple, thus producing a different equilibrium and non-equilibrium behaviour. Surprisingly, to the best of our knowledge, there is no in-depth study of this case. For this reason, more study is necessary.

On the other hand, in a nonquilibrium many-body system, a more intriguing question is about its thermalization or relaxation property. Thermalization is among the most fundamental processes for a many-body interacting system. All systems do not thermalize. An example is the Fermi-Pasta-Ulam paradox [ ] which shows confinement in phase-space so that the system stay non-thermalized for a long time. A lot of research on thermalization property has been done in other classical system. Ref.\cite{Tsuchiya2000}\cite{Yawn1997} studied the one dimensional gravitational system with Lyapunov exponents. Ref.\cite{Jin2013} studied the ring of harmonic oscillators and magnetic moment by examine the canonical distribution of a subsystem . In our paper, we will also probe the system's thermalization time scales from the perspective of canonical distribution as well as Lyapunov exponent. %With these approaches, we will show that the thermalization time scale manifest a competition between total energy and interaction energy, so there is not any singular behavior in the thermalization time scales except the trivial divergence at zero interaction.

This paper is organized as follows: In Section \ref{section:preparation}, we perform a scaling to reduce the number of parameters of the system. In addition, we briefly observe the ground state and reveal two dynamically distinct regime of the system, i.e. strong interaction regime and weak interaction regime. In Section \ref{section:breathing frequency}, we evaluate the breathing frequency according to its mechanism, and compare that with our numerical simulation result. Besides, we also introduce a rotating frame method of treating this problem, which is a convenient tool to analyze simple harmonic systems. In section \ref{section:Thermalization}, we focus on the thermalization property. We first examine the canonical distribution of energy to find out the threshold of thermalization, i.e. under what parameter the thermalization time scale is very large. In the end of this section, we measure the Lyapunov Exponent to quantify the time scales and confirm the former discussion. Then we point out that there are two kind of thermalization behavior that could be measured in this system, one is about energy distribution while another is about shape of cloud in phase-space. Since here we find two sense of ``thermalizing", we will redefine our terms -- When a system's energy distribution goes to equilibrium, we still call it ``thermalized".  But when a system's distribution in phase-space goes to equilibrium, we call it ``relaxed" to distiguish it from ``thermalized". So in section \ref{section:Shape} we actually study the ``thermalizing" of the shape of the cloud in phase-space, or the ``relaxation" behavior of the system. There we discovered a long-lasting mode, which turns out to be a quadruple mode in phase-space.

\section{Preparation}\label{section:preparation}
\subsection{Scaling behaviour}
We investigated the classical gas in harmonic trap with a repulsive interaction. We chose one of the simplest types of interaction:
\begin{equation}
F=\left\lbrace
\begin{aligned}
&F=F_0 	&,x\textless\sigma\\
&F=0 	&,x\textgreater\sigma
\end{aligned}
\right.
\label{eq:preparation1}
\end{equation}
so the Hamiltonian is
\begin{equation}
H=\frac{1}{2}m\omega_0^2\sum_{i}x_i^2+\frac{1}{2}m\sum_i v_i^2+\sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}

Using the transformation Eq.\ref{eq:transform}, we can rewrite Eq.\ref{eq:preparation2} into the following form:
\begin{equation}
\tilde{H}=\frac{1}{2}\sum_{i}\tilde{x_i}^2+\frac{1}{2}\sum_i\left(\frac{d\tilde{x}}{d\tilde{t}}\right)^2+\sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless 1}F_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)
\end{equation}
\begin{equation}
\left\lbrace
\begin{aligned}
&\tilde{H}&=&\frac{H}{m\omega_0^2\sigma^2}\\
&\tilde{x_i}&=&\frac{x_i}{\sigma}\\
&\tilde{t}&=&\omega_0t\\
&\tilde{F_0}&=&\frac{F_0}{m\omega_0^2\sigma}\\
\end{aligned}
\right.
\label{eq:transform}
\end{equation}

In this manner, we reduced the number of the parameters in our model to three: $\tilde{H}$, $\tilde{F_0}$, $N$. Equivalently, we set $\sigma$, $\omega_0$ and $m$ to 1 in our numerical simulation. Throughout, we will use ``E" and ``$F_0$" to denote their reduced version $\tilde{H}$, $\tilde{F_0}$.

@@

\subsubsection{Ground states}
For the ground state of this system, one could make the following prediction: \\(1) When the $\sigma$ is larger than the size of the cloud (in view of the scaling behavior, large $\sigma$ is equivalent to small $F_0$ limit when we fix $\sigma=1$), it can be easily shown that the ground state is a solid-like state, where each particle keeps an equal distance ($\frac{2F_0}{\omega_0^2}$) with its nearest neighbors . The low-lying excitation is particles' independent harmonic oscillation around their equilibrium position.\\(2) When $F_0$ is very large the molecules act like a hardcore gas, while the size of particle is just $\sigma$.

In the Fig.\ref{fig:GS1}, we can see this crossover from the ``solid-like" limit (left) to the ``hardcore gas" limit (right).

@@

\section{Breathing Frequency}\label{section:breathing frequency}
Usually, to see the response of quantum gas to a quench, one begin simulation with the ground state and then suddenly vary some parameter by a little, for instance, $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$. In that case, one will observe the radius R of the cloud oscillating at a certain frequency, which is usually called ``breathing mode".
However, it is quite different in the classical version. Firstly, it is no longer necessary to focus on ground state. In quantum case, the reason why we care about ground state and low lying excitations is that BEC occur at low temperature. But in classical case, there is no BEC. So it is not necessary to focus on low lying excitations in classical case. Besides, since classical limit is $\hbar\rightarrow 0$, no matter how low the kT is, $\hbar/kT$ is always zero so even in low energy regime, the classical behaviors could be quite different from the quantum one. Secondly, it is no longer necessary to vary parameters ``by a little" in the quench. In quantum case, the small quench $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$ means the spectrum of Hamiltonian is shifted slightly, thus one may expect to see some beat phenomenon({\color{red}{?}}). But in classical case, since $\hbar$ vanishes, no matter how small the quench is, the reason that cause beats in quantum case no longer work in classical context. Thus it is not necessary to limit ourselves to small quench.

Let's consider quenching the harmonic potential, i.e., $\omega_0\rightarrow\omega_0'=\omega_0+\Delta\omega_0$. We notice that X and P (scaled) are symmetric for simple harmonic oscillator. In weak interaction limit, the distribution cloud should be circular when the gas is completely thermalized. The difference between thermalized cloud before and after quench is that they are circular under different scaling ratio of P and X axis. If we look at both of them in the phase-space rescaled according to $\omega_0'$, we will find shape of the cloud before the quench is an ellipse while the final one is a circle.

Since the number of particles we use in the Molecular Dynamics simulation is limited (5 to 50), the random noise is significant. If we start with a state whose configuration in phase-space is an ellipse slightly deviated from circle, the oscillating amplitude will be too small to be distinguished from noise. Therefore, we start from an extreme case -- line distribution (random x, zero p) and observe the oscillation of R(t). 
\begin{figure}[hbtp]
\center
\includegraphics[scale=0.32]{ZhiyuPictures/freq_scanF_differentE_log_2_with_oscillation_demo.png}
\caption{breathing mode frequency measured at different E and $F_0$. Insert: a demonstration of oscillation, $E_{trap}$ is the total potential energy of particles in the trap, which is proportional to $\sum_{is}{x_i^2}$, thus manifest the oscillation behavior of radius R(t) defined in Eq.\ref{eq:def_of_R}. The breathing frequency is measured by taking Fourier transformation of R(t).}
\label{fig:Breathingfrequency1}
\end{figure}

@@

\subsection{Rotating frame in Phase-space}
If there is a non-circular distribution cloud in phase-space, e.g., a line-shape cloud in Fig.\ref{fig:Breathingfrequency2_0}, driven by the harmonic trap, the cloud will rotate at frequency $\omega_0$. When there is interaction, we may think there is a small modification of the frequency $\omega=\omega_0+\delta$. Now we want to understand how $\delta$ is dependent on parameters, so it is better to stand in the rotating frame in the phase-space. In this frame, the picture we will see is as follow:\\
\begin{enumerate}[\textbf{*}]
\item All particles are stationary when there is no interaction.
\item The real X and P axis are rotating counter-clockwise, using real X axis to measure distance between particles to determine the interaction at each moment.
\item When there is interaction, particles will gain a ``velocity" in phase space: $\dot{P}=F/m$ . Other effects are all cancelled by frame rotation.
\end{enumerate}

This rotating frame picture is actually a classical version of the interaction picture, which turns out to symplifly our dynamic analysis. The stationary frame and rotating frame is shown in Fig.\ref{fig:Breathingfrequency2_1}.

In our system, we choose initial velocity to be zero, which means, the particles are aligned on $X|_{t=0}$ axis in phase-space at the beginning. If the breathing frequency of the system is $\omega=\omega_0+\delta$, we will expect to see a line (which may gradually deform into an oval cloud or even an isotropic cloud) rotating at $\delta$ in rotating frame. Since the motion we see is an additional rotation in rotating frame, we will call it precession. This precession directly leads to the $\delta$ shift in breathing frequency. Fig.\ref{fig:Breathingfrequency2_1} show this precession motion in the rotating frame in phase-space.

\begin{figure*}
\subfigure[$t=0$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=1-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=1-5T_rev.eps}
\end{minipage}
}
\subfigure[$t=\frac{1}{5}T$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=2-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=2-5T_rev.eps}
\end{minipage}
}\subfigure[$t=\frac{2}{5}T$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=3-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=3-5T_rev.eps}
\end{minipage}
}\subfigure[$t=\frac{3}{5}T$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=4-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=4-5T_rev.eps}
\end{minipage}
}\subfigure[$t=\frac{4}{5}T$]{
\begin{minipage}[b]{0.18\linewidth}
\includegraphics[scale=0.21]{ZhiyuPictures/stationary_frame_t=5-5T_rev.eps} 
\includegraphics[scale=0.22]{ZhiyuPictures/rotating_frame_t=5-5T_rev.eps}
\end{minipage}
}
\caption{N=50,F0=100,E=50000. Upper(lower) five pictures are snapshot of the configuration of the cloud in the stationary(rotating) frame at t=$\frac{1}{5}T$, $\frac{2}{5}T$, $\frac{3}{5}T$, $\frac{4}{5}T$, $T$ respectively.}
\label{fig:Breathingfrequency2_0}
\end{figure*}

@@

(See Fig.\ref{fig:Breathingfrequency3}(a))Two particles flip to another line(thick gray line) which deviate a small angle $\alpha$ away from original configuration. Similar process happen when there are more particles. In every period of harmonic oscillation, each particle meet N particles and collide 2N times, while half of the collisions (N times) are between particles with huge difference in momentum, which is just the process shown in Fig.\ref{fig:Breathingfrequency3}. (As for the remaining half of collisions, colliding particles have small difference in their momentum. Let us say, both particles have positive momentum. In Fig.\ref{fig:Breathingfrequency3}, it can be shown that two particles on the same half of the P-axis seldom contribute to the rotation of distribution.)

Let us estimate the precession angle $\alpha$ in one collision event by $\frac{\sigma}{R}$, which follows from the two-particle case study. Radius of the cloud R can be estimated according to $E=Nm\omega_0^2R^2$
We will get the precession angular velocity 
\begin{equation}
\delta=2\omega_0\frac{N^\frac{3}{2}m^{\frac{1}{2}}E^{-\frac{1}{2}}\sigma\omega_0}{2\pi}
\label{eq:breathingfrequency1}
\end{equation}

When $F_0$ is not that large, the frequency behavior seems complicated (see Fig.\ref{fig:Breathingfrequency1}). Usually, when $F_0$ increase, frequency goes down below 2 first and then rise up and converge to some value higher than 2. The reason could also be well understood with the mechanism described before. In the discussion above, we go to high $F_0$ limit, which means particles exchange momentum in infinitesimal time. For $F_0$ is not very big, the finite interaction time should be taken into consideration. During this time, the motion of each particle in phase-space is exactly moving along the direction of the real P axis at 'velocity' $F0/m$. Since the real P axis is rotating counter-clockwise, the particle will follow a circular trajectory(orange and green dashed line in Fig.\ref{fig:Breathingfrequency3}) the $\alpha$ that we once use to estimate the precession motion should decrease with the increase of interaction time. There should be some ``critical point" where $\alpha$ goes from plus to minus (dashed green line). It is easy to see that the contribution of each collision event to the phase-space precession of the cloud is negative in Fig.\ref{fig:Breathingfrequency3})(b). So at this point, we will see the precession  slows down or even be reversed, which means the breathing frequency is 2 or a little bit lower than 2.

@@

The dependence of frequency on energy and $F_0$ is shown in Fig.\ref{fig:Breathingfrequency4}. For the first thing, as expected when $F_0$ is high the frequency behavior is well-described by linear prediction Eq.\ref{eq:breathingfrequency1}. Secondly, the figure shows the frequency's dependence on $F_0$. As we explained, when $F_0$ is not so large, the precession angle decrease. In language of real space, the finite $F_0$ create a finite approaching time, thus consumes the time saved by finite-range collision. As a result, the frequency should be lower when $F_0$ decrease, and reach 2 earlier.

the critical point that frequency decreases to 2 can be estimate by:
\begin{equation}
\omega_0 \tau=\alpha
\end{equation}
which means that the approaching time $\tau$ (estimated by $R/F_0$) exactly cancels the precession angle we derived before. One can show that this condition is equivalent to
\begin{equation}
\frac{E}{N}=F_0\sigma
\end{equation}
which can also be understood as the criterion of whether two particles will bounce or pass by if one estimate the average internal energy of a pair with $\frac{E}{N}$.

Before completing the discussion, it is necessary to point out that none of the argument here require ``energy thermalization". In addition, if the system is completely thermalized in terms of their phase angles in phase-space, the argument will no longer work -- there is a stable isotropic distribution in phase space, no matter how the cloud rotate, no oscillation can be observed.

@@

We can consider their motion in their center of mass frame ``C". In frame C, one can easily show that when there is no interaction what we will see is that two particles bound together by a harmonic trap centered at C (black line in Fig.\ref{fig:thermalization2}). So their relative motion is also a harmonic oscillation.

In this case, the system has two frequency components. The first one is the frequency of C, which is just $\omega_0$ . The other one is the frequency of their relative motion $\omega_r$, which is slightly deviated from $\omega_0$. The deviation grows with the increase of $F_0$ and the decrease of internal energy of the pair. As a result, when $F_0$ is small or when E is large, a beat with low frequency $\omega_r-\omega_0=\delta$ will exist. This phenomenon means energy transfer between particles are inefficient, thus the thermalization should be slow.

@@

\subsection{Thermalization condition}
The main obstruct to get an thermalized state is the low frequency beat mentioned before. As a result, to achieve ergodic state, we have to avoid such low frequency oscillation. According to our discussion before, the solution is to make internal energy of each two-particle pair not ``too large". Though it is impossible to express the internal energy of every pair in terms of the total energy $E$, we can estimate it by the average energy, which is $\frac{E}{N}$. At least they are of the same order. The thermalization condition could be given by:
\begin{equation}
F_0\sigma\sim\frac{E}{N}
\label{eq:thermalizatiton condition}
\end{equation}

@@

As is shown in Fig.\ref{fig:thermalization5} above, the critical point for reaching ``Boltzmann-like" distribution (later we will verify that it is Boltzmann distribution) is $F_0=500$-$F_0=1000$ (we always set $\sigma=1$), while the average energy $\sim 1000$, as we expected, they are of the same order. As a supplementary proof of our argument, Fig.\ref{fig:thermalization6} and Fig.\ref{fig:thermalization7} show the great difference of single-particle energy between two examples of below and above the thermalization threshold ($F_0=200$ $\&$ $F_0=1000$ for $N=10, E=10000$): in the former one there is always some particle maintained at high excitation, while in the latter the system probably goes to ergodic. To date, we have verified that the threshold of thermalization, which is $\frac{NF_0\sigma}{E}\sim1 $ , and corresponding feature of their energy distribution in two side of the threshold.
